Supplementary Information

### 2020-09-16 writing/transect_paper.Rmd
Additional changes for MER and ENA submissions

*Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
*A. bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair (`r Biostrings::readDNAStringSet(here::here("data/clusters/ITS2.fasta.gz")) %>% magrittr::extract(startsWith(names(.), "f2ab49e8")) %>% Biostrings::width() %>% unique() %>% divide_by(1, .) %>% subtract(1, .) %>% percent(digits = 1, format = "f")` similarity in ITS2).
The abundance of the second ASV was `r readd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~. > 0) %>% percent(format = "f", digits = 1) %>% paste(collapse = " and ")` that of the primary *A. bisporus* ASV in the two Illumina positive controls.
The consistency of this ratio across replicate positive controls suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
Despite higher total sequencing depth, this ASV was not identified from the Ion Torrent dataset.

@@

 - Sequence data, including raw reads and ASV sequences, will be deposited at the European Nucleotide Archive (ENA) prior to final publication.
 - Nucleotide alignment and ML tree will be deposited at Dryad prior to final publication [@furneauxdata2020].
 - R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
These packages are currently being prepared for submission to Bioconductor.
If they are not accepted at Bioconductor prior to final publication, they will be archived at Dryad.
 - FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate. The versions used in this paper will be archived at Dryad prior to publication.
 - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect.

>>>>

*Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
*A. bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair (`r Biostrings::readDNAStringSet(here::here("data/clusters/ITS2.fasta.gz")) %>% magrittr::extract(startsWith(names(.), "f2ab49e8")) %>% Biostrings::width() %>% unique() %>% divide_by(1, .) %>% subtract(1, .) %>% percent(digits = 1, format = "f")` similarity in ITS2).
The abundance of the second ASV was `r readd(agaricus_reads) %>% as.data.frame() %>% filter(rowSums(.) > 0) %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~. > 0) %>% percent(format = "f", digits = 1) %>% paste(collapse = " and ")` that of the primary *A. bisporus* ASV in the two Illumina positive controls.
The consistency of this ratio across replicate positive controls suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
Despite higher total sequencing depth, this ASV was not identified from the Ion Torrent dataset.

@@

 - Trimmed, demultiplexed sequencing reads have been deposited at the European Nucleotide Archive (ERA) under Project accession number PRJEB37385. Accession numbers are given in Supplementary Files 1 and 2.
 - Consensus ASV sequences will also be deposited at ENA prior to publication.
 - Nucleotide alignment and ML tree will be deposited at Dryad prior to final publication [@furneauxdata2020].
 - R packages `LSUx`, `tzara`, `phylotax`, and `FUNGuildR` are available on Github at https://github.com/brendanf/LSUx, https://github.com/brendanf/tzara, https://github.com/brendanf/phylotax, and https://github.com/brendanf/FUNGuildR.
These packages are currently being prepared for submission to CRAN/Bioconductor.
If they are not accepted prior to final publication, snapshots will be archived at Dryad.
 - FASTA-format files for the RDP, Warcup, and Unite reference databases with unified classifications, as well as scripts used to generate them, are available at https://github.com/brendanf/reannotate. The versions used in this paper will be archived at Dryad prior to publication.
 - Bioinformatics pipeline and analysis scripts are available at https://github.com/oueme-fungi/oueme-fungi-transect.

### 2020-09-16 writing/transect_supplement.Rmd
Additional changes for MER and ENA submissions

**Supplementary file 1**: Mapping between samples and indexed primers for short amplicons.

**Supplementary file 2**: Mapping between samples and indexed primers for long amplicons.

@@

```{r variofit, results="asis"}
 readd(variofit_table) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high),
    "-",
    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

>>>>

**Supplementary file 1**: Mapping between samples, indexed primers, and read accession numbers for short amplicons.

**Supplementary file 2**: Mapping between samples, indexed primers, and read accession numbers for long amplicons.

@@

```{r variofit, results="asis"}
 readd(variofit_table) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high) | mantel == FALSE,
    "-",
    sprintf(
      "%s (%s–%s)",
      formatC(estimate, digits = 2, format = "fg"),
      formatC(conf.low, digits = 2, format = "fg"),
      formatC(conf.high, digits = 2, format = "fg")
      )
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(
    names_from = "term",
    values_from = "value",
    id_cols = c("guild", "amplicon", "tech", "metric", "algorithm"),
    values_fill = list(value = "-")
  ) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```
